---
title: "ISU Monarch Data Summary for NCIG"
author: "Jarad Niemi and Seth Appelgate"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{rotating}
  - \usepackage{caption} 
  - \captionsetup[table]{skip=10pt}
output: 
  pdf_document:
    fig_caption: TRUE
---

```{r setup, purl=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(echo    = FALSE, 
               results = 'asis',
               message = FALSE)
```

```{r packages}
suppressWarnings(library("tidyverse"))
library("ISUmonarch")
library("xtable")
options(xtable.comment = FALSE)
```

\setlength{\tabcolsep}{3pt}


```{r sites}
sites = as.character(unique(transect$transectID))
sites18 = as.character(unique(sites[!(sites %in% c("tuth1a", "tuth2a", "tsie1a"))]))
sites = as.character(unique(sites[!(sites %in% c("tuth1a", "tuth2a", "tsie1a", "tfis1a"))]))

milkweed_sites = c("tarm1b", "tall2a", "tber3a", "tnkn1a")


ramet_species = unique(ramet$`Nectar Plant Species`)[!(unique(ramet$`Nectar Plant Species`) %in% "milkweed strip: common milkweed ramet")]
milkweed_species = unique(ramet$`Nectar Plant Species`)[(unique(ramet$`Nectar Plant Species`) %in% "milkweed strip: common milkweed ramet")]
```


Table \ref{t:ramet2017} provides mean count of milkweed ramets averaged across all 
data collection events. 
Values in table represent the specific counting unit for each species, not necessarily a whole plant.
Table \ref{t:density2017ramet} provides mean density of milkweed ramets (per 100m2) averaged across all data collection events.
Values in table represent the specific counting unit for each species, not necessarily a whole plant.

<!-- Table \ref{t:environment} provides a binary response of whether there were any  -->
<!-- flowering plants or milkweed present in the area around the site during any  -->
<!-- round. -->


<!-- Ramet -->

```{r ramet_table_function}
ramet_table <- function(yr, flt, transects, label) {
  ramet %>%
    
    mutate(round = as.factor(round)) %>%
    
    filter(transectID %in% transects, year == yr) %>%
    
    group_by(`Nectar Plant Species`, transectID, round) %>%
    summarize(count = sum(count)) %>%
    complete(transectID, round, fill = list(count = 0)) %>%
    group_by(`Nectar Plant Species`, transectID) %>%
    summarize(total = mean(count) %>% round(1)) %>%
    complete(transectID = transects) %>%
    replace(.==0, NA) %>%
    tidyr::spread(transectID, total, fill=NA, drop=FALSE) %>%
    arrange(`Nectar Plant Species`) %>% 
    
    filter(`Nectar Plant Species` %in% flt) %>%
    
    xtable(digits = 1,
           caption=paste(yr, "ramet plant species: mean count across all surveys"),
           label = paste("t:",label,sep="")) %>%
    print(include.rownames = FALSE, NA.string="-", 
          table.placement = "p",
          caption.placement = "top",
          size="\\tiny",
          floating.environment = "sidewaystable") 
}
```

```{r ramet, warning=FALSE}
ramet_table(yr=2017, flt=ramet_species, transects=sites, label = "ramet2017")
ramet_table(yr=2018, flt=ramet_species, transects=sites18, label = "ramet2018")
```


```{r milkweed, warning=FALSE}
ramet_table(yr=2017, flt=milkweed_species, transects=milkweed_sites, label="milkweed2017")
ramet_table(yr=2018, flt=milkweed_species, transects=milkweed_sites, label="milkweed2018")
```



\newpage



```{r ramet_density_table}
ramet_density_table <- function(yr, flt, transects, label) {
ramet %>%
  mutate(`Nectar Plant Species` = 
           fct_recode(`Nectar Plant Species`,
                      smartweed = "swamp smartweed",
                      smartweed = "pennsylvania smartweed",
                      `culvers root` = "culver's root",
                      `shepherds purse` = "shepherd's purse",
                      `yellow sweet clover` = "yellow sweetclover"),
         year = as.character(year),
         round = as.factor(round)) %>% 
  
  filter(transectID %in% transects, year == yr) %>%
  
  left_join(subset(ISUmonarch::survey, year==yr) %>% 
              mutate(round = as.factor(round)) %>%
              select(transectID, length, round), by=c("transectID", "round")) %>%
  
  group_by(`Nectar Plant Species`, transectID, round) %>%
  summarize(density = sum(count, na.rm=TRUE)/as.numeric(unique(length))) %>%
  complete(transectID, round, fill = list(density = 0)) %>%
  
  group_by(`Nectar Plant Species`, transectID) %>%
  summarize(density = mean(density, na.rm=TRUE)) %>%
  
  complete(transectID = transects) %>%
  ungroup() %>% 
  select(`Nectar Plant Species`, transectID, density) %>%
  replace(.==0, NA) %>%
  tidyr::spread(transectID, density, fill=NA) %>%
  arrange(`Nectar Plant Species`) %>% 
  
  filter(`Nectar Plant Species` %in% flt) %>%
    
  xtable(digits = 2,
           caption=paste(yr, "Nectar plant species density: average density (count / m2) across all rounds"),
           label=paste("t:",label,sep="")) %>%
    print(include.rownames = FALSE, NA.string="-", 
          table.placement = "p",
          caption.placement = "top",
          size="\\tiny",
          floating.environment = "sidewaystable") 
}
```

```{r ramet_density, warning=FALSE, message=FALSE}
ramet_density_table(yr=2017, flt=ramet_species, transects=sites,   label="density2017ramet")
ramet_density_table(yr=2018, flt=ramet_species, transects=sites18, label="density2018ramet")
```


```{r milkweed_density, warning=FALSE, message=FALSE}
ramet_density_table(yr=2017, flt=milkweed_species, transects=milkweed_sites,   label="density2017milkweed")
ramet_density_table(yr=2018, flt=milkweed_species, transects=milkweed_sites, label="density2018milkweed")
```


```{r nectar_density_heatmap}
ramet_density_heatmap <- function(yr, flt, transects) {
  tmp = ramet %>%
    mutate(`Nectar Plant Species` = fct_recode(`Nectar Plant Species`,
                                               smartweed = "swamp smartweed",
                                               smartweed = "pennsylvania smartweed",
                                              `culvers root` = "culver's root",
                                              `shepherds purse` = "shepherd's purse",
                                              `yellow sweet clover` = "yellow sweetclover"),
           year = as.character(year),
           round = as.factor(round)) %>% 
    
    filter(transectID %in% transects, year == yr) %>%
  
    left_join(subset(ISUmonarch::survey, year==yr) %>% 
                mutate(round = as.factor(round)) %>%
                select(transectID, length, round), by=c("transectID", "round")) %>%
    
    group_by(`Nectar Plant Species`, transectID, round) %>%
    summarize(density = sum(count, na.rm=TRUE)/as.numeric(unique(length))) %>%
    complete(transectID, round, fill = list(density = 0)) %>%
  
    group_by(`Nectar Plant Species`, transectID) %>%
    summarize(density = mean(density, na.rm=TRUE)) %>%
  
    complete(transectID = transects) %>%
    ungroup() %>% 
    select(`Nectar Plant Species`, transectID, density) %>%
    replace(.==0, NA) %>%
    tidyr::spread(transectID, density, fill=NA) %>%
    arrange(`Nectar Plant Species`) %>% 
  
    filter(`Nectar Plant Species` %in% flt) %>%
  
    gather(variable, value, -`Nectar Plant Species`) %>%
  
    mutate(presence = ifelse(is.na(value), 0, 1))
    
    
    tran.order = tmp %>% group_by(variable) %>% summarise(avg.density = mean(value, na.rm=TRUE)) %>% arrange(desc(avg.density)) %>% select(variable) %>% pull(variable)
  
  # species.order = tmp %>% group_by(`Nectar Plant Species`) %>% summarise(avg.density = mean(value, na.rm=TRUE)) %>% arrange(avg.density) %>% select(`Nectar Plant Species`) %>% pull(`Nectar Plant Species`)
  
  species.order = tmp %>% group_by(`Nectar Plant Species`) %>% summarise(avg.density = mean(presence, na.rm=TRUE)) %>% arrange(avg.density) %>% select(`Nectar Plant Species`) %>% pull(`Nectar Plant Species`)

  tmp$variable <- factor(tmp$variable, levels = tran.order)
  tmp$`Nectar Plant Species` <- factor(tmp$`Nectar Plant Species`, levels = species.order)
  
  ggplot(tmp, aes(x=variable, y=`Nectar Plant Species`, z= value)) + 
  geom_tile(aes(fill = value), color="grey") + 
  theme_bw() + labs(x="", y="", fill="Log(Density)") +
  scale_fill_gradient(na.value = 'white', low="navy", high="salmon", trans="log") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + 
  scale_x_discrete(position = "top") +
  theme(axis.text.x.top = element_text(vjust = 0.5))
}
```


```{r ramet_heatmap_17, warning=FALSE, message=FALSE, fig.height=3, fig.width=9, fig.cap="This is a heatmap for density of ramet species with transects on the columns and species on the rows sorting transects by average density and sorting species by average presence for year 2017."}
ramet_density_heatmap(yr=2017, flt=ramet_species, transects=sites)
```


```{r ramet_heatmap_18, warning=FALSE, message=FALSE, fig.height=3, fig.width=9, fig.cap="This is a heatmap for density of ramet species with transects on the columns and species on the rows sorting transects by average density and sorting species by average presence for year 2018."}
ramet_density_heatmap(yr=2018, flt=ramet_species, transects=sites18)
```


```{r milkweed_heatmap_17, warning=FALSE, message=FALSE, fig.height=2, fig.width=9, fig.cap="This is a heatmap for density of milkweed species with transects on the columns and species on the rows sorting transects by average density and sorting species by average presence for year 2017."}
ramet_density_heatmap(yr=2017, flt=milkweed_species, transects=milkweed_sites)
```


```{r milkweed_heatmap_18, warning=FALSE, message=FALSE, fig.height=2, fig.width=9, fig.cap="This is a heatmap for density of milkweed species with transects on the columns and species on the rows sorting transects by average density and sorting species by average presence for year 2018."}
ramet_density_heatmap(yr=2018, flt=milkweed_species, transects=milkweed_sites)
```

