---
title: "ISU Monarch Data Summary for NCIG"
author: "Jarad Niemi, Seth Appelgate, and Nehemias Ulloa"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{rotating}
  - \usepackage{caption} 
  - \captionsetup[table]{skip=10pt}
output: 
  pdf_document:
    fig_caption: TRUE
---

```{r setup, purl=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(echo    = FALSE, 
               results = 'asis',
               message = FALSE)
```

```{r packages}
suppressWarnings(library("tidyverse"))
library("ISUmonarch")
library("xtable")
options(xtable.comment = FALSE)
```

\setlength{\tabcolsep}{3pt}


```{r sites}
NCIG_sites = as.character(unique(transect$transectID))
NCIG_sites = as.character(unique(NCIG_sites[!(NCIG_sites %in% c("tuth1a", "tuth2a", "tsie1a"))]))
```


Table \ref{t:monarch} provides total number of monarchs sighted anywhere around 
site during entire time at site, number of monarchs sighted during a survey 
period, number of milkweed ramets (an upper limit of 20 per round per site, 
although the same milkweed may have been surveyed in different rounds), 
number of monarch eggs counted on milkweed plants (if no milkweed present, 
no monarch eggs present, number of monarch larvae (caterpillars) counted on 
milkweed plants (if no milkweed present, no larvae present).


<!-- Monarch -->

```{r monarch, warning=FALSE}
monarch_summary <- monarch %>% 
  filter(transectID %in% NCIG_sites) %>%
  
  mutate(group = "survey",
         group = ifelse(grepl("instar",distance),                 "larvae", group),
         group = ifelse(grepl("butterfly_[0-9]instar",distance),  "butterfly_larvae", group),
         group = ifelse(grepl("common_[0-9]instar",distance),     "common_larvae", group),
         group = ifelse(grepl("swamp_[0-9]instar",distance),      "swamp_larvae", group),
         group = ifelse(distance == "stems",                      "ramets" , group),
         group = ifelse(grepl("ramet", distance),                 "ramets" , group),
         group = ifelse(grepl("butterfly_ramet",distance),        "butterfly_ramets", group),
         group = ifelse(grepl("common_ramet",distance),           "common_ramets", group),
         group = ifelse(grepl("swamp_ramet",distance),            "swamp_ramets", group),
         group = ifelse(grepl("eggs", distance),                  "eggs"  , group),
         group = ifelse(grepl("butterfly_eggs",distance),         "butterfly_eggs", group),
         group = ifelse(grepl("common_eggs",distance),            "common_eggs", group),
         group = ifelse(grepl("swamp_eggs",distance),             "swamp_eggs", group),
         group = ifelse(distance == "extra_monarchs",             "extra",  group)) %>%
  
  select(transectID, year, group, count) %>%
  group_by(transectID, year, group) %>%
  summarize(count = sum(count)) %>%
  tidyr::spread(group, count, fill=0) %>%
  mutate(Total  = survey + extra) %>%
  ungroup() %>%
  magrittr::set_colnames(tools::toTitleCase(gsub("_", " ", colnames(.)))) %>%
  select(-Extra) %>%
  # select(year, transectID, Total, Survey, Ramets, Eggs, Larvae) %>%
  
  tidyr::gather(Monarch, count, -Year, -transectID) %>%
  tidyr::spread(transectID, count, fill = NA) %>%
  
  mutate(Monarch = factor(Monarch, 
                          levels = c("Total","Survey",
                                     "Ramets","Butterfly Ramets","Common Ramets","Swamp Ramets",
                                     "Eggs","Butterfly Eggs","Common Eggs","Swamp Eggs",
                                     "Larvae","Butterfly Larvae","Common Larvae","Swamp Larvae"
                                     ))) %>%
  
  arrange(Monarch,Year)
  
monarch_summary %>%
  xtable(digits = 0,
         caption="Monarch adults (total and during survey), milkweed ramets, monarch eggs on milkweed, and monarch larvae on milkweed. Note ramets, eggs, and larvae were collected differently for 2016.",
         label = "t:monarch") %>%
  print(include.rownames = FALSE, NA.string="-",
        caption.placement = "top",
        size="\\small",
        floating.environment = "sidewaystable",
        hline.after = c(-1,seq(from=0, to=(length(unique(monarch_summary$year))*5), by=3))) 
```






```{r monarch_plot_1, warning=FALSE, messages=FALSE, fig.height=9, fig.width=9, fig.cap="This plots the total number of monarchs sighted anywhere around the site during entire time at site. The colors correspond to the different trasects. \\label{fig:monarch_plot_1}"}
monarch %>% 
  filter(transectID %in% NCIG_sites) %>%
  
  mutate(group = "survey",
         group = ifelse(grepl("instar",distance),     "instar", group),
         group = ifelse(distance == "stems",          "ramets" , group),
         group = ifelse(grepl("ramet", distance),     "ramets" , group),
         group = ifelse(grepl("eggs", distance),           "eggs"  , group),
         group = ifelse(distance == "extra_monarchs", "extra",  group)) %>%
  
  select(transectID, year, group, count) %>%
  group_by(transectID, year, group) %>%
  summarize(count = sum(count)) %>%
  tidyr::spread(group, count, fill=0) %>%
  mutate(Total  = survey + extra) %>%
  ungroup() %>%
  rename(Survey = survey,
         Ramets = ramets,
         Eggs = eggs,
         Larvae = instar) %>%
  select(year, transectID, Total, Survey, Ramets, Eggs, Larvae) %>%
  
  tidyr::gather(Monarch, count, -year, -transectID) %>%
  left_join(transect, by=c("transectID")) %>%
  mutate(year  = factor(year, levels=paste0(min(year):max(year)))) %>%
  ggplot(data=.) + 
    geom_point(aes(x=year, y=count, color=transectID)) +
    geom_line( aes(x=year, y=count, color=transectID, group=transectID)) +
    facet_grid(grant~Monarch, scales="free_y") + 
    theme_bw() + 
    labs(x="Year", y="Count", color="Transect")
```



