---
title: "ISU Monarch Data Summary for NCIG"
author: "Jarad Niemi, Seth Appelgate, and Nehemias Ulloa"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{rotating}
  - \usepackage{caption} 
  - \captionsetup[table]{skip=10pt}
output: 
  pdf_document:
    fig_caption: TRUE
---

```{r setup, purl=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(echo    = FALSE, 
               results = 'asis',
               message = FALSE)
```

```{r packages}
suppressWarnings(library("tidyverse"))
library("ISUmonarch")
library("xtable")
options(xtable.comment = FALSE)
```

\setlength{\tabcolsep}{3pt}


```{r species, warning=FALSE, message=FALSE}
species = unique(nectar$`Nectar Plant Species`)
species = species[order(species)]
species = species %>% data.frame(common_name=.) %>% 
  left_join(ISUmonarch::species, by="common_name")

plntd_species = species %>% filter(planted_non_planted=="planted") %>% pull(common_name)
nonplntd_species = species %>% filter(planted_non_planted=="non planted") %>% pull(common_name)

nonplntd_species1 = nonplntd_species[split(1:length(nonplntd_species), cut(1:length(nonplntd_species), 2, labels = FALSE))[[1]]]
nonplntd_species2 = nonplntd_species[split(1:length(nonplntd_species), cut(1:length(nonplntd_species), 2, labels = FALSE))[[2]]]

all_species = species %>% pull(common_name)
```



```{r sites}
every_transect = ISUmonarch::transect %>% 
  filter(!(transectID %in% c("tuth1a", "tuth2a", "tsie1a"))) %>%
  select(transectID, grant) %>%
  arrange(transectID)

ncig  = every_transect %>% filter(grant=="ncig") %>% pull(transectID) %>% as.character
pork  = every_transect %>% filter(grant=="pork") %>% pull(transectID) %>% as.character
iacig = every_transect %>% filter(grant=="iacig") %>% pull(transectID) %>% as.character
soy   = every_transect %>% filter(grant=="soy") %>% pull(transectID) %>% as.character
all_transects = every_transect %>% pull(transectID) %>% as.character



make_align_string <- function(x, brk) {
  k <- ncol(x)
  format_str <- ifelse(sapply(x, is.numeric), "r", "l")
  if(is.null(brk)){
    paste0(c("r", format_str), collapse = "")
  }else{
    paste0(c("r", format_str[c(1:brk)], "|", format_str[-c(1:brk)]), collapse = "")
  }
}

```


Table \ref{t:nectar2016} and \ref{t:nectar2017} provide mean count of nectar plant species averaged across all 
data collection events. 
Values in table represent the specific counting unit for each species, not necessarily a whole plant.
Table \ref{t:density2016} and \ref{t:density2017} provide nectar plant species density (per 100m2) averaged across all data collection events. 



<!-- Nectar -->

```{r nectar_table_function}
nectar_table <- function(yr, flt, transects, label) {
  nectar %>%
    
    mutate(round = as.factor(round)) %>%
    
    filter(transectID %in% transects, year == yr) %>%
    
    group_by(`Nectar Plant Species`, transectID, round) %>%
    summarize(count = sum(count)) %>%
    complete(transectID, round, fill = list(count = 0)) %>%
    group_by(`Nectar Plant Species`, transectID) %>%
    summarize(total = mean(count) %>% round(1)) %>%
    complete(transectID = transects) %>%
    replace(.==0, NA) %>%
    tidyr::spread(transectID, total, fill=NA, drop=FALSE) %>%
    arrange(`Nectar Plant Species`) %>% 
    
    filter(`Nectar Plant Species` %in% flt) %>%
    
    xtable(digits = 1,
           caption=paste(yr, "nectar plant species: mean count across all surveys"),
           label = paste("t:",label,sep="")) %>%
    print(include.rownames = FALSE, 
          NA.string="-", 
          # rotate.colnames = TRUE,
          table.placement = "p",
          caption.placement = "top",
          size="\\tiny",
          floating.environment = "sidewaystable") 
}
```

```{r nectar, warning=FALSE, eval=FALSE}
# 2016 NCIG
nectar_table(yr=2016, 
             flt=native_species,
             transects = ncig,
             label = "nectar2016_ncig_native")
nectar_table(yr=2016, 
             flt=intro_species,
             transects = ncig,
             label = "nectar2016_ncig_intro")
nectar_table(yr=2016, 
             flt=both_species,
             transects = ncig,
             label = "nectar2016_ncig_both")
# 2016 Pork
nectar_table(yr=2016, 
             flt=native_species,
             transects = pork,
             label = "nectar2016_pork_native")
nectar_table(yr=2016, 
             flt=intro_species,
             transects = pork,
             label = "nectar2016_pork_intro")
nectar_table(yr=2016, 
             flt=both_species,
             transects = pork,
             label = "nectar2016_pork_both")
# 2016 ICIG
nectar_table(yr=2016, 
             flt=native_species,
             transects = iacig,
             label = "nectar2016_iacig_native")
nectar_table(yr=2016, 
             flt=intro_species,
             transects = iacig,
             label = "nectar2016_iacig_intro")
nectar_table(yr=2016, 
             flt=both_species,
             transects = iacig,
             label = "nectar2016_iacig_both")
# 2016 Soy
nectar_table(yr=2016, 
             flt=native_species,
             transects = soy,
             label = "nectar2016_soy_native")
nectar_table(yr=2016, 
             flt=intro_species,
             transects = soy,
             label = "nectar2016_soy_intro")
nectar_table(yr=2016, 
             flt=both_species,
             transects = soy,
             label = "nectar2016_soy_both")


# 2017 NCIG
nectar_table(yr=2017, 
             flt=native_species,
             transects = ncig,
             label = "nectar2017_ncig_native")
nectar_table(yr=2017, 
             flt=intro_species,
             transects = ncig,
             label = "nectar2017_ncig_intro")
nectar_table(yr=2017, 
             flt=both_species,
             transects = ncig,
             label = "nectar2017_ncig_both")
# 2017 Pork
nectar_table(yr=2017, 
             flt=native_species,
             transects = pork,
             label = "nectar2017_pork_native")
nectar_table(yr=2017, 
             flt=intro_species,
             transects = pork,
             label = "nectar2017_pork_intro")
nectar_table(yr=2017, 
             flt=both_species,
             transects = pork,
             label = "nectar2017_pork_both")
# 2017 ICIG
nectar_table(yr=2017, 
             flt=native_species,
             transects = iacig,
             label = "nectar2017_iacig_native")
nectar_table(yr=2017, 
             flt=intro_species,
             transects = iacig,
             label = "nectar2017_iacig_intro")
nectar_table(yr=2017, 
             flt=both_species,
             transects = iacig,
             label = "nectar2017_iacig_both")
# 2017 Soy
nectar_table(yr=2017, 
             flt=native_species,
             transects = soy,
             label = "nectar2017_soy_native")
nectar_table(yr=2017, 
             flt=intro_species,
             transects = soy,
             label = "nectar2017_soy_intro")
nectar_table(yr=2017, 
             flt=both_species,
             transects = soy,
             label = "nectar2017_soy_both")


# 2018 NCIG
nectar_table(yr=2018, 
             flt=native_species,
             transects = ncig,
             label = "nectar2018_ncig_native")
nectar_table(yr=2018, 
             flt=intro_species,
             transects = ncig,
             label = "nectar2018_ncig_intro")
nectar_table(yr=2018, 
             flt=both_species,
             transects = ncig,
             label = "nectar2018_ncig_both")
# 2018 Pork
nectar_table(yr=2018, 
             flt=native_species,
             transects = pork,
             label = "nectar2018_pork_native")
nectar_table(yr=2018, 
             flt=intro_species,
             transects = pork,
             label = "nectar2018_pork_intro")
nectar_table(yr=2018, 
             flt=both_species,
             transects = pork,
             label = "nectar2018_pork_both")
# 2018 ICIG
nectar_table(yr=2018, 
             flt=native_species,
             transects = iacig,
             label = "nectar2018_iacig_native")
nectar_table(yr=2018, 
             flt=intro_species,
             transects = iacig,
             label = "nectar2018_iacig_intro")
nectar_table(yr=2018, 
             flt=both_species,
             transects = iacig,
             label = "nectar2018_iacig_both")
# 2018 Soy
nectar_table(yr=2018, 
             flt=native_species,
             transects = soy,
             label = "nectar2018_soy_native")
nectar_table(yr=2018, 
             flt=intro_species,
             transects = soy,
             label = "nectar2018_soy_intro")
nectar_table(yr=2018, 
             flt=both_species,
             transects = soy,
             label = "nectar2018_soy_both")
```



```{r nectar_density_table}
nectar_density_table <- function(yr, flt, transects, label, brk=NULL, capsites) {
nectar %>%
  mutate(`Nectar Plant Species` = 
           fct_recode(`Nectar Plant Species`,
                      smartweed = "swamp smartweed",
                      smartweed = "pennsylvania smartweed",
                      `culvers root` = "culver's root",
                      `shepherds purse` = "shepherd's purse",
                      `yellow sweet clover` = "yellow sweetclover"),
         year = as.character(year),
         round = as.factor(round)) %>% 
  
  filter(transectID %in% transects, year == yr) %>%
  
  left_join(subset(ISUmonarch::survey, year==yr) %>% 
              mutate(round = as.factor(round)) %>%
              select(transectID, length, round), by=c("transectID", "round")) %>%
  
  group_by(`Nectar Plant Species`, transectID, round) %>%
  summarize(density = sum(count, na.rm=TRUE)/as.numeric(unique(length))) %>%
  complete(transectID, round, fill = list(density = 0)) %>%
  
  group_by(`Nectar Plant Species`, transectID) %>%
  summarize(density = mean(density, na.rm=TRUE)) %>%
  
  complete(transectID = transects) %>%
  ungroup() %>% 
  select(`Nectar Plant Species`, transectID, density) %>%
  replace(.==0, NA) %>%
  tidyr::spread(transectID, density, fill=NA) %>%
  arrange(`Nectar Plant Species`) %>% 
  
  filter(`Nectar Plant Species` %in% flt) %>%
    
  xtable(digits = 2,
         caption=paste(yr, "Nectar plant species density: average density (count / m2) across all rounds for", capsites, "grant transects"),
         label=paste("t:",label,sep=""),
         align = make_align_string(., brk)) %>%
    print(include.rownames = FALSE, 
          NA.string="-", 
          # rotate.colnames = TRUE,
          table.placement = "p",
          caption.placement = "top",
          size="\\small",
          floating.environment = "sidewaystable",
          sanitize.colnames.function = function(x){
            rotated <- paste("\\begin{sideways}", x, "\\end{sideways}")
            # put all column names into sideways environments for the rotation.
            return(c("Nectar Plant Species", paste(rotated[-1], collapse="&")))}
            # replaces first column name with something else (not rotated).
    )
}
```

```{r nectar_density, warning=FALSE, message=FALSE}
# 2016 NCIG
nectar_density_table(yr=2016, 
             flt=plntd_species,
             transects = ncig,
             capsites = "NCIG",
             label = "density2016_ncig_native")
nectar_density_table(yr=2016, 
             flt=nonplntd_species1,
             transects = ncig,
             capsites = "NCIG",
             label = "density2016_ncig_intro")
nectar_density_table(yr=2016, 
             flt=nonplntd_species2,
             transects = ncig,
             capsites = "NCIG",
             label = "density2016_ncig_both")
# 2016 Pork & Soy
nectar_density_table(yr=2016, 
             flt=plntd_species,
             transects = c(pork, soy),
             capsites = "Pork and Soy",
             label = "density2016_pork_native",
             brk = length(pork))
nectar_density_table(yr=2016, 
             flt=nonplntd_species1,
             transects = c(pork, soy),
             capsites = "Pork and Soy",
             label = "density2016_pork_intro",
             brk = length(pork))
nectar_density_table(yr=2016, 
             flt=nonplntd_species2,
             transects = c(pork, soy),
             capsites = "Pork and Soy",
             label = "density2016_pork_both",
             brk = length(pork))
# 2016 ICIG
nectar_density_table(yr=2016, 
             flt=plntd_species,
             transects = iacig,
             capsites = "IACIG",
             label = "density2016_iacig_native")
nectar_density_table(yr=2016, 
             flt=nonplntd_species1,
             transects = iacig,
             capsites = "IACIG",
             label = "density2016_iacig_intro")
nectar_density_table(yr=2016, 
             flt=nonplntd_species2,
             transects = iacig,
             capsites = "IACIG",
             label = "density2016_iacig_both")



# 2017 NCIG
nectar_density_table(yr=2017, 
             flt=plntd_species,
             transects = ncig,
             capsites = "NCIG",
             label = "density2017_ncig_native")
nectar_density_table(yr=2017, 
             flt=nonplntd_species1,
             transects = ncig,
             capsites = "NCIG",
             label = "density2017_ncig_intro")
nectar_density_table(yr=2017, 
             flt=nonplntd_species2,
             transects = ncig,
             capsites = "NCIG",
             label = "density2017_ncig_both")
# 2017 Pork & Soy
nectar_density_table(yr=2017, 
             flt=plntd_species,
             transects = c(pork, soy),
             capsites = "Pork and Soy",
             label = "density2017_pork_native",
             brk = length(pork))
nectar_density_table(yr=2017, 
             flt=nonplntd_species1,
             transects = c(pork, soy),
             capsites = "Pork and Soy",
             label = "density2017_pork_intro",
             brk = length(pork))
nectar_density_table(yr=2017, 
             flt=nonplntd_species2,
             transects = c(pork, soy),
             capsites = "Pork and Soy",
             label = "density2017_pork_both",
             brk = length(pork))

# 2017 ICIG
nectar_density_table(yr=2017, 
             flt=plntd_species,
             transects = iacig,
             capsites = "IACIG",
             label = "density2017_iacig_native")
nectar_density_table(yr=2017, 
             flt=nonplntd_species1,
             transects = iacig,
             capsites = "IACIG",
             label = "density2017_iacig_intro")
nectar_density_table(yr=2017, 
             flt=nonplntd_species2,
             transects = iacig,
             capsites = "IACIG",
             label = "density2017_iacig_both")



# 2018 NCIG
nectar_density_table(yr=2018, 
             flt=plntd_species,
             transects = ncig,
             capsites = "NCIG",
             label = "density2018_ncig_native")
nectar_density_table(yr=2018, 
             flt=nonplntd_species1,
             transects = ncig,
             capsites = "NCIG",
             label = "density2018_ncig_intro")
nectar_density_table(yr=2018, 
             flt=nonplntd_species2,
             transects = ncig,
             capsites = "NCIG",
             label = "density2018_ncig_both")
# 2018 Pork & Soy
nectar_density_table(yr=2018, 
             flt=plntd_species,
             transects = c(pork, soy),
             capsites = "Pork and Soy",
             label = "density2018_pork_native",
             brk = length(pork))
nectar_density_table(yr=2018, 
             flt=nonplntd_species1,
             transects = c(pork, soy),
             capsites = "Pork and Soy",
             label = "density2018_pork_intro",
             brk = length(pork))
nectar_density_table(yr=2018, 
             flt=nonplntd_species2,
             transects = c(pork, soy),
             capsites = "Pork and Soy",
             label = "density2018_pork_both",
             brk = length(pork))
# 2018 ICIG
nectar_density_table(yr=2018, 
             flt=plntd_species,
             transects = iacig,
             capsites = "IACIG",
             label = "density2018_iacig_native")
nectar_density_table(yr=2018, 
             flt=nonplntd_species1,
             transects = iacig,
             capsites = "IACIG",
             label = "density2018_iacig_intro")
nectar_density_table(yr=2018, 
             flt=nonplntd_species2,
             transects = iacig,
             capsites = "IACIG",
             label = "density2018_iacig_both")

```


```{r nectar_density_heatmap}
nectar_density_heatmap <- function(yr, flt, transects, plnt="planted") {

tmp=nectar %>%
  mutate(year = as.character(year),
         round = as.factor(round)) %>% 
  
  left_join(ISUmonarch::species[,c("common_name", "planted_non_planted")], by=c("Nectar Plant Species"="common_name")) %>%
    
  filter(transectID %in% transects, year %in% yr, planted_non_planted %in% plnt) %>%
  
  left_join(subset(ISUmonarch::survey, year==yr) %>% 
              mutate(round = as.factor(round)) %>%
              select(transectID, length, round), by=c("transectID", "round")) %>%
  
  group_by(`Nectar Plant Species`, transectID, round) %>%
  summarize(density = sum(count, na.rm=TRUE)/as.numeric(unique(length))) %>%
  complete(transectID, round, fill = list(density = 0)) %>%
  
  group_by(`Nectar Plant Species`, transectID) %>%
  summarize(density = mean(density, na.rm=TRUE)) %>%
  
  complete(transectID = transects) %>%
  ungroup() %>% 
  select(`Nectar Plant Species`, transectID, density) %>%
  replace(.==0, NA) %>%
  tidyr::spread(transectID, density, fill=NA) %>%
  arrange(`Nectar Plant Species`) %>% 
  
  filter(`Nectar Plant Species` %in% flt) %>%
  
  gather(variable, value, -`Nectar Plant Species`) %>%
  
  mutate(presence = ifelse(is.na(value), 0, 1))
  
  tran.order = tmp %>% group_by(variable) %>% summarise(avg.density = mean(value, na.rm=TRUE)) %>% arrange(desc(avg.density)) %>% select(variable) %>% pull(variable)
  
  # species.order = tmp %>% group_by(`Nectar Plant Species`) %>% summarise(avg.density = mean(value, na.rm=TRUE)) %>% arrange(avg.density) %>% select(`Nectar Plant Species`) %>% pull(`Nectar Plant Species`)
  
  species.order = tmp %>% group_by(`Nectar Plant Species`) %>% summarise(avg.density = mean(presence, na.rm=TRUE)) %>% arrange(avg.density) %>% select(`Nectar Plant Species`) %>% pull(`Nectar Plant Species`)

  tmp$variable <- factor(tmp$variable, levels = tran.order)
  tmp$`Nectar Plant Species` <- factor(tmp$`Nectar Plant Species`, levels = species.order)
  
  ggplot(tmp, aes(x=variable, y=`Nectar Plant Species`, z= value)) + 
  geom_tile(aes(fill = value), color="grey") + 
  theme_bw() + labs(x="", y="", fill="Log(Density)") +
  scale_fill_gradient(na.value = 'white', low="navy", high="salmon", trans="log") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + 
  scale_x_discrete(position = "top") +
  theme(axis.text.x.top = element_text(vjust = 0.5))
}
```

Here are some heatmaps for density of species planted with transects on the columns and species on the rows sorting transects by average density and sorting species by average presence.


```{r nectar_heatmap_16, warning=FALSE, message=FALSE, fig.height=9, fig.width=9, fig.cap="This plot is a heatmap for density of species planted and non-planted with transects on the columns and species on the rows sorting transects by average density and sorting species by avg presence for year 2016. Caption 1234"}
nectar_density_heatmap(yr = 2016, 
                       flt = all_species, 
                       transects = all_transects, 
                       plnt=c("planted", "non planted"))
```


```{r nectar_heatmap_17, warning=FALSE, message=FALSE, fig.height=9, fig.width=9, fig.cap="This plot is a heatmap for density of species planted and non-planted with transects on the columns and species on the rows sorting transects by average density and sorting species by avg presence for year 2017. Caption 1235"}
nectar_density_heatmap(yr = 2017, 
                       flt = all_species, 
                       transects = all_transects, 
                       plnt=c("planted", "non planted"))
```


```{r nectar_heatmap_18, warning=FALSE, message=FALSE, fig.height=9, fig.width=9, fig.cap="This plot is a heatmap for density of species planted and non-planted with transects on the columns and species on the rows sorting transects by average density and sorting species by avg presence for year 2018. Caption 1236"}
nectar_density_heatmap(yr = 2018, 
                       flt = all_species, 
                       transects = all_transects[!(all_transects %in% c("tuth1a","tuth2a","tsie1a","tfis1a"))],
                       plnt=c("planted", "non planted"))
```




```{r nectar_heatmap2_planted_16, warning=FALSE, message=FALSE, fig.height=5, fig.width=9, fig.cap="This plot is a heatmap for density of species planted with transects on the columns and species on the rows sorting transects by average density and sorting species by avg presence for year 2016. Caption 1237"}
nectar_density_heatmap(yr = 2016, 
                       flt = all_species, 
                       transects = all_transects,
                       plnt=c("planted"))
```


```{r nectar_heatmap2_planted_17, warning=FALSE, message=FALSE, fig.height=6, fig.width=9, fig.cap="This plot is a heatmap for density of species planted with transects on the columns and species on the rows sorting transects by average density and sorting species by avg presence for year 2017. Caption 1238"}
nectar_density_heatmap(yr = 2017, 
                       flt = all_species, 
                       transects = all_transects,
                       plnt=c("planted"))
```


```{r nectar_heatmap2_planted_18, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.cap="This plot is a heatmap for density of species planted with transects on the columns and species on the rows sorting transects by average density and sorting species by average presence for year 2018. Caption 1239"}
nectar_density_heatmap(yr = 2018,
                       flt = all_species, 
                       transects = all_transects[!(all_transects %in% c("tuth1a","tuth2a","tsie1a","tfis1a"))],
                       plnt=c("planted"))
```


```{r nectar_heatmap2_non_planted_16, warning=FALSE, message=FALSE, fig.height=9, fig.width=9, fig.cap="This plot is a heatmap for density of species non planted with transects on the columns and species on the rows sorting transects by average density and sorting species by average presence for year 2016. Caption 1240"}
nectar_density_heatmap(yr = 2016, 
                       flt = all_species, 
                       transects = all_transects,
                       plnt=c("non planted"))
```


```{r nectar_heatmap2_non_planted_17, warning=FALSE, message=FALSE, fig.height=9, fig.width=9, fig.cap="This plot is a heatmap for density of species non planted with transects on the columns and species on the rows sorting transects by average density and sorting species by average presence for year 2017. Caption 1241"}
nectar_density_heatmap(yr = 2017, 
                       flt = all_species, 
                       transects = all_transects,
                       plnt=c("non planted"))
```


```{r nectar_heatmap2_non_planted_18, warning=FALSE, message=FALSE, fig.height=9, fig.width=9, fig.cap="This plot is a heatmap for density of species non planted with transects on the columns and species on the rows sorting transects by average density and sorting species by average presence for year 2018. Caption 1242"}
nectar_density_heatmap(yr = 2018, 
                       flt = all_species, 
                       transects = all_transects[!(all_transects %in% c("tuth1a","tuth2a","tsie1a","tfis1a"))],
                       plnt=c("non planted"))
```





```{r inflorescence, echo=FALSE, warning=FALSE}
d <- nectar %>%
  mutate(ymd = lubridate::ymd(paste(year,month,day,sep="-"))) %>%
  group_by(`Nectar Plant Species`, year, transectID) %>%
  summarize(inflorescence = sum(count),
            richness = sum(count>0)) %>%
  mutate(richness = ifelse(richness>0, 1, 0)) %>%
  filter(year!="2017" | transectID!="tall1a" | `Nectar Plant Species`!="prostrate vervain") %>%
  group_by(year, transectID) %>%
  summarize(inflorescence = sum(inflorescence),
            richness = sum(richness)) %>%
  left_join(ISUmonarch::transect, by="transectID") %>%
  filter(!(transectID %in% c("tuth1a", "tuth2a", "tsie1a"))) %>%
  ungroup() %>%
  mutate(year  = factor(year, levels=paste0(min(year):max(year))))
```


```{r inflorescence_plot, fig.height=9, fig.width=9, fig.cap="This plot shows the inflorescence for the nectar data across years and faceting the transects by grant excluding tall1a in 2017 for ncig sites. Caption 1111 \\label{fig:inflorescence}"}
# ggplot(d, aes(x=year, y=inflorescence)) +
#   geom_point(aes(color=transectID)) +
#   geom_line(aes(color=transectID, group=transectID)) +
#   facet_wrap(~grant) +
#   theme_bw() + 
#   scale_y_continuous(trans = "log") +
#   labs(y="log(Inflorescence)", x="Year")


ggplot(d, aes(x=year, y=inflorescence)) +
  geom_point(aes(color=transectID)) +
  geom_line(aes(color=transectID, group=transectID)) +
  facet_wrap(~grant) +
  theme_bw() + 
  labs(y="Inflorescence", x="Year")
```


```{r richness_plot, fig.height=9, fig.width=9, fig.cap="This plot shows the richness for the nectar data across years and faceting the transects by grant. Caption 1122 \\label{fig:richness}"}
# ggplot(d, aes(x=year, y=log(richness))) +
#   geom_point(aes(color=transectID)) +
#   geom_line(aes(color=transectID, group=transectID)) +
#   facet_wrap(~grant) +
#   theme_bw() +
#   labs(y="log(Richness)", x="Year")

ggplot(d, aes(x=year, y=richness)) +
  geom_point(aes(color=transectID)) +
  geom_line(aes(color=transectID, group=transectID)) +
  facet_wrap(~grant) +
  theme_bw() +
  labs(y="Richness", x="Year")
```




```{r anova_data, warning=FALSE, message=FALSE}
da <- nectar %>%
  mutate(ymd = lubridate::ymd(paste(year,month,day,sep="-")),
         year = as.factor(as.character(year))) %>%
  group_by(`Nectar Plant Species`, year, transectID, round) %>%
  
  left_join(subset(ISUmonarch::survey) %>% 
              mutate(round = as.factor(round),
                     year = as.factor(year)) %>%
              select(transectID, length, round, year), by=c("transectID", "round", "year")) %>%
  
  summarize(inflorescence = sum(count),
            richness = sum(count>0),
            density = sum(count, na.rm=TRUE)/as.numeric(unique(length))) %>%
  
  mutate(richness = ifelse(richness>0, 1, 0)) %>%
  complete(transectID, round, fill = list(density = 0)) %>%
  
  filter(year!="2017" | transectID!="tall1a" | `Nectar Plant Species`!="prostrate vervain") %>%
  
  left_join(ISUmonarch::transect, by="transectID") %>%
  filter(!(transectID %in% c("tuth1a", "tuth2a", "tsie1a"))) %>%
  
  left_join(ISUmonarch::species, by = c("Nectar Plant Species" = "common_name")) %>%
  ungroup() %>%
  mutate(year = as.numeric(as.character(year)),
         year  = factor(year, levels=paste0(min(year):max(year))),
         transectID = as.factor(transectID),
         grant = as.factor(grant),
         round = as.factor(round)
         )
```


```{r anova_m1, warning=FALSE, message=FALSE}
# options(contrasts = c("contr.sum","contr.poly"))
model <- lm(richness ~ round + planted_non_planted + year + grant, data=da)
xtable(drop1(model, .~., test="F"))
```



```{r anova_m2, warning=FALSE, message=FALSE}
model2 <- lm(density ~ round + planted_non_planted + year + grant, data=da)
xtable(drop1(model2, .~., test="F"))
```






